---
title: "GARCH Modeling"
output:
  html_document:
    df_print: paged
---

## Load Required Libraries
```{r}
# install.packages(c("quantmod", "zoo", "ggplot2", "tseries", "rugarch",
#                    "bayesGARCH", "stats", "timeDate", "timeSeries",
#                    "forecast", "xts", "FinTS", "strucchange"),
#                 dependencies = TRUE,
#                 repos = "https://cloud.r-project.org/"
# )

suppressPackageStartupMessages({
  library(quantmod) # for getSymbols
  library(zoo) # for na.omit
  library(ggplot2) # for plotting
  library(tseries) # for adf.test
  library(rugarch) # for GARCH modeling
  library(bayesGARCH) # for Bayesian GARCH
  library(stats) # for spectrum, acf, pacf
  library(timeDate) # for business days
  library(timeSeries) # for holidaySIFMA
  library(forecast) # for auto.arima
  library(xts) # for xts objects
  library(FinTS) # for ArchTest
  library(strucchange)
})
```

## Load and Plot 10-Year Treasury Yield Data
```{r} 
# "Market Yield on U.S. Treasury Securities at 10-Year Constant Maturity,
# Quoted on an Investment Basis"

range_date_from <- "2010-01-01" # start date for data
range_date_to <- format(Sys.Date() - 7, "%Y-%m-%d") # end date - last week
print(paste("Loading DGS10 data from", range_date_from, "to", range_date_to))

#getSymbols("DGS10", src = "FRED", from = range_date_from, to = range_date_to) # until yesterday
DGS10 <- read.csv("DGS10.csv")
DGS10 <- xts(DGS10$DGS10, order.by = as.Date(DGS10$observation_date))
plot(DGS10, main = "DGS10 - 10Y Treasury Constant Maturity Rate",
     col = "blue", ylab = "Yield (%)", xlab = "Date")
```

The yield curve clearly shows three regimes: post-GFC low rates, 
pandemic crash (~2020), and sharp rise post-2022 (inflation, Fed tightening).
Some missing values are visible.

## Preliminary Data Analysis
### Check for Missing Values
```{r}
# Helper functions
get_year_as_num <- function(date) as.numeric(format(as.Date(date), "%Y"))
is_bizday <- function(dates) {
  years <- get_year_as_num(range_date_from):get_year_as_num(range_date_to)
  holidays <- as.Date(holidayNYSE(years))
  # Columbus Day, Veterans Day, National Day of Mourning
  extra_holidays <-as.Date(
      c("2010-10-11", "2010-11-11", "2011-10-10", "2011-11-11", "2012-10-08",
        "2012-11-12", "2013-10-14", "2013-11-11", "2014-10-13", "2014-11-11",
        "2015-10-12", "2015-11-11", "2016-10-10", "2016-11-11", "2017-10-09",
        "2018-10-08", "2018-11-12", "2018-12-05", "2019-10-14", "2019-11-11",
        "2020-10-12", "2020-11-11", "2021-10-11", "2021-11-11", "2022-10-10",
        "2022-11-11", "2023-10-09", "2024-10-14", "2024-11-11"))
  holidays <- timeDate(sort(unique(c(holidays, extra_holidays))))
  isBizday(dates, holidays)
}

# analysis of all days in the range
temp_time_seq <- timeSequence(from = range_date_from,
                              to = range_date_to,
                              by = "day")
all_dates <- as.Date(temp_time_seq)
business_days <- as.Date(temp_time_seq[is_bizday(temp_time_seq)])
non_business_days <- as.Date(setdiff(all_dates, business_days))
earliest_date <- min(all_dates)
latest_date <- max(all_dates)

# dates in DGS10
days_in_data <- index(DGS10)
duplicate_days_in_data <- days_in_data[duplicated(days_in_data)]
business_days_in_data <- business_days[business_days %in% days_in_data]
non_business_days_in_data <- as.Date(non_business_days[non_business_days %in% days_in_data])
earliest_date_in_data <- min(days_in_data)
latest_date_in_data <- max(days_in_data)
real_days_in_data <- days_in_data[days_in_data %in% all_dates]
missing_days <- as.Date(setdiff(all_dates, days_in_data))
missing_business_days <- as.Date(setdiff(business_days, days_in_data))
missing_non_business_days <- as.Date(setdiff(non_business_days, days_in_data))

# Analysis of NAs
days_with_na_in_data <- index(DGS10[is.na(DGS10)])
business_days_with_na_in_data <- business_days[business_days %in% days_with_na_in_data]
non_business_days_with_na <- non_business_days[non_business_days %in% days_with_na_in_data]

# printing results
cat("Business Days in Data:", length(business_days_in_data), "\n")
cat("Non-Business Days in Data:", length(non_business_days_in_data), "\n")
cat("Missing Days in Data:", length(missing_days), "\n")
cat("Missing Business Days in Data:", length(missing_business_days),
    paste(missing_business_days, collapse = ", "), "\n")
cat("Missing Non-Business Days in Data:", length(missing_non_business_days), "\n")
print("\n")
cat("Days with NA in Data:", length(days_with_na_in_data), "\n")
cat("Business Days with NA in Data:", length(business_days_with_na_in_data), "\n")
cat("Non-Business Days with NA in Data:", length(non_business_days_with_na), "\n")
print("\n")
cat("Business Days with NA in Data:", paste(business_days_with_na_in_data, collapse = ", "), "\n")
```

--> Missing dates in DGS10 are all non-business days (eg holidays) <br>
--> Weirdly, some non-business days are in the data, however with NA values <br>
--> All NAs are on non-business days, so we can safely remove them, meaning no values are missing <br>

### Removing NAs
```{r}
DGS10 <- na.omit(DGS10)
plot(DGS10, main = "DGS10 - 10Y Treasury Constant Maturity Rate",
     col = "blue", ylab = "Yield (%)", xlab = "Date")
```

### Check for Stationarity
```{r}
# testing if its stationary
adf.test(DGS10, alternative = "stationary")
```

p-value is very large, so we have to accept H0 --> data is not stationary <br>
-------------
We use the squared daily changes to model volatility.
```{r}
# diff computes the daily change. The first value is NA since there is no
# previous value to subtract from the first one, so we remove it.
DGS10_diff <- diff(DGS10)[-1]
DGS10_diff_sq <- DGS10_diff^2
colnames(DGS10_diff_sq) <- "SquaredDailyChange"
plot(DGS10_diff_sq, main = "Squared Daily Yield Changes in DGS10",
     col = "darkred", ylab = "∆ Yield (%)", xlab = "Date")
```

Major spikes around early 2020 (pandemic) and post-2022 (inflation uncertainty)

GARCH models require weakly stationary time series which can be tested with ADF test.
```{r}
# testing stationarity of the daily changes
adf.test(DGS10_diff_sq, alternative = "stationary")
```

p-value is very low, so we can say that the data is weakly stationary now (there is enough statistical evidence to reject H0). 

### Frequency Spectrum of Daily Changes
```{r}
# Frequency spectrum of daily changes
spectrum(DGS10_diff_sq, method = "pgram",
         main = "Frequency Spectrum of DGS10 Squared Daily Changes",
         col = "darkgreen", ylab = "Power in dB", xlab = "Frequency",
         log = "dB")
```

```{r}
spectrum(DGS10_diff_sq, method = "pgram",
         main = "Frequency Spectrum of DGS10 Squared Daily Changes",
         col = "darkgreen", ylab = "Power", xlab = "Frequency",
         log = "no")
```

The spectrum is relatively flat with higher power at very low frequencies.
The spike near 0 indicates long-memory or persistent behavior in volatility.
The high power at low frequencies implies low-frequency (long-term) cycles dominate.

### ACF and PACF of Daily Changes
```{r}
acf(DGS10_diff_sq, main = "ACF of DGS10 Squared Daily Changes")
```

The ACF decays slowly and remains statistically significant for many lags.
Reinforces the presence of volatility clustering.
Suggests that past squared returns can predict future volatility to some extent.

```{r}
pacf(DGS10_diff_sq, main = "PACF of DGS10 Squared Daily Changes")
```

The PACF shows significant spikes up to lag ~7, especially lags 1–5.
This suggests that the squared daily changes exhibit short-term persistence or clustering.
There is conditional heteroskedasticity, supporting the use of ARCH/GARCH models for volatility modeling.

### ARCH-Test
```{r}
# ARCH-LM test for conditional heteroskedasticity
for (L in c(1, 7, 14, 30, 60, 120, 365*2, 365*3, 365*4)) {
  print(ArchTest(DGS10_diff, lags = L, demean = TRUE))
}
```

Volatility in 10-year Treasury yields shows strong short- and medium-term persistence, and statistically significant ARCH effects for up to 3 years. Beyond that, the influence of past squared shocks on current variance becomes negligible.

## GARCH Model Specification and Fitting
### Mean model selection

```{r}
acf(DGS10_diff, main = "ACF of DGS10 Daily Changes")
```

```{r}
pacf(DGS10_diff_sq, main = "PACF of DGS10 Daily Changes")
```



```{r}
auto.arima(coredata(DGS10_diff),
           d = 0, # already differenced
           max.p = 10, max.q = 10,
           stationary = TRUE,
           seasonal = FALSE)
```

The mean of the differenced yield series (i.e. daily yield changes) is not statistically different from zero.
--> mean model: ARIMA(0,0,0) with zero mean
The ACF also supports this selection because we don't see any significant autocorrelation there.

### Model Fitting
We should use the simple differenced data (not the squared values). The squared resiudals are implicitly handled inside the GARCH model.

```{r}
# standard GARCH(1,1) model with Student-t innovations
# on the first difference of the 10-year Treasury yield series
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  # mean model with no ARMA terms, just a constant mean
  mean.model = list(armaOrder = c(0, 0), include.mean = FALSE),
  # Student-t distribution, assumed distribution of shocks/random errors
  distribution.model = "std"
)

fit <- ugarchfit(spec, DGS10_diff)
show(fit)
```
### GARCH Model Evaluation Summary

- MLE converged because log-likelihood is finite
- all parameters have low p-values, so they are all important
- $\alpha + \beta = 0.050496 + 0.941409 = 0.991905 < 1$, so it's still stationary but there is strong volatility persistence
- ARCH LM Test: p > 0.7 --> no remaining ARCH effects
- Stability Test: model parameters are stable
- Ljung-Box Test: no autocorrelation in residuals or squared residuals
- Sign Bias: Tests whether residual sign alone affects conditional variance
- Negative Sign Bias: Tests whether negative shocks increase volatility more than expected, p = 0.19 --> small assymetry
- Positive Sign Bias: Tests whether positive shocks increase volatility differently
- Joint Effect: F-test of all three biases combined

### GARCH model selection with grid search
Testing GARCH(p, q, distribution) with parameters $p, q \in \{1, ..., 5\}$, and distribution for normal and student-t.

```{r}
best_aic <- Inf
best_spec <- NULL
best_fit <- NULL
distributions <- c("norm", "std")

for (dist in distributions){
  for (p in 1:5) {
    for (q in 1:5) {
      spec <- ugarchspec(
        variance.model = list(model = "sGARCH", garchOrder = c(p, q)),
        mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
        distribution.model = dist
      )
      fit <- tryCatch(ugarchfit(spec, DGS10_diff, solver = "hybrid"), error = function(e) NULL)
      
      if (!is.null(fit)) {
        aic <- infocriteria(fit)[1]  # [1] = AIC
        if (aic < best_aic) {
          best_aic <- aic
          best_spec <- spec
          best_fit <- fit
        }
      }
    }
  }
}

show(best_fit)

```

### Inspecting the fit of the model - standardized residual diagnostics

```{r}
spec_norm <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0)),
  distribution.model = "std"
)
fit_norm <- ugarchfit(spec_norm, DGS10_diff)

resids <- residuals(fit_norm, standardize = TRUE)

hist(resids, breaks = 50, freq = FALSE, col = "lightblue",
     main = "Standardized Residuals",
     xlab = "Value")

# Add standard normal density curve
curve(dnorm(x, mean = 0, sd = 1), col = "black", lwd = 2, add = TRUE)

# Add legend
legend("topright", legend = "Standard Normal", 
       col = "black", lwd = 2, bty = "n")
```

```{r}
acf(resids, main = "ACF of GARCH(1,1) Standardized Residuals")
```

```{r}
acf(resids^2, main = "ACF of GARCH(1,1) Squared Standardized Residuals")
```

```{r}
plot(resids, type = "l", main = "Standardized Residuals", 
     ylab = "Residual", xlab = "Time", col = "darkred")
```


Best parameter setup (based on lowest AIC) is: GARCH(1, 1) with student-t innovations. The standardized residuals approximate a standard normal distribution very well which means that the student-t innovations are appropriate for GARCH.

Also, the selected model is backed up by the results: the standardized residuals are approximately standard normal, there is no autocorrelation between either the residuals or the squared residuals. The residuals seem to be more or less homoscedastic - at least there is less volatility clustering. 




### GARCH with MLE, rolling prediction for the last 3 months

```{r}
# Length of time series
N <- length(DGS10_diff)

# Set forecast length to ~3 months (63 trading days)
forecast_horizon <- 63

# Define train/test
train <- DGS10_diff[1:(N - forecast_horizon)]
test <- DGS10_diff[(N - forecast_horizon + 1):N]

# GARCH(1,1) specification with Student-t errors
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "std"
)

# Rolling 1-day ahead forecast
roll <- ugarchroll(
  spec,
  data = DGS10_diff,
  forecast.length = forecast_horizon,
  refit.every = 1,
  refit.window = "moving",
  solver = "hybrid",
  calculate.VaR = FALSE
)

# Extract forecasted sigma (volatility)
forecasted_sigma <- as.numeric(roll@forecast$density[, "Sigma"])
forecasted_var <- forecasted_sigma^2  # conditional variance

# Use squared returns as proxy for realized volatility
realized_returns <- tail(DGS10_diff, forecast_horizon)
realized_vol <- realized_returns^2

# Data for plotting
comparison_df <- data.frame(
  Day = 1:forecast_horizon,
  ForecastedVar = forecasted_var,
  RealizedVar = realized_vol
)

# Plot
ggplot(comparison_df, aes(x = Day)) +
  geom_line(aes(y = ForecastedVar), color = "blue", size = 1.2) +
  geom_line(aes(y = RealizedVar), color = "red", linetype = "dashed", size = 1.2) +
  labs(
    title = "GARCH Forecasted Conditional Variance vs Squared Realized Returns",
    y = "Variance", x = "Forecast Day"
  ) +
  theme_minimal()
```

3 months prediction: GARCH catches the trends but reacts slowly to them.

### GARCH With Indicators

```{r}
# Setup
N <- length(DGS10_diff)
forecast_horizon <- 63
full_index <- index(DGS10_diff)
realized_returns <- tail(DGS10_diff, forecast_horizon)
realized_var <- realized_returns^2

# Define rolling predictions container
forecasted_var_with_indicators <- numeric(forecast_horizon)

# Create full event matrix matching the entire series
event_indicators_full <- data.frame(
  COVID = ifelse(full_index >= as.Date("2020-03-01") & full_index <= as.Date("2020-05-15"), 1, 0),
  EnergyFed2022 = ifelse(full_index >= as.Date("2022-03-01") & full_index <= as.Date("2022-10-01"), 1, 0),
  PreElection = ifelse(full_index >= as.Date("2023-12-01") & full_index <= as.Date("2024-06-15"), 1, 0)
)

# Rolling forecast with external regressors
for (i in 1:forecast_horizon) {
  
  # Subset data up to time t + i - 1
  train_slice <- DGS10_diff[1:(N - forecast_horizon + i - 1)]
  reg_slice <- event_indicators_full[1:(N - forecast_horizon + i - 1), ]

  # Take the matching regressors for variance equation
  spec_step <- ugarchspec(
    variance.model = list(model = "sGARCH", garchOrder = c(1, 1),
                          external.regressors = as.matrix(reg_slice)),
    mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
    distribution.model = "std"
  )

  # Fit and forecast
  fit_step <- ugarchfit(spec_step, train_slice, solver = "hybrid", solver.control = list(trace = 0))
  fc <- ugarchforecast(fit_step, n.ahead = 1)
  forecasted_var_with_indicators[i] <- sigma(fc)^2
}

```




```{r}
# Comparison dataframe
comparison_df <- data.frame(
  Day = 1:forecast_horizon,
  RealizedVar = realized_var,
  ForecastedVar_NoIndicators = forecasted_var,  # from ugarchroll
  ForecastedVar_Indicators = forecasted_var_with_indicators
)

ggplot(comparison_df, aes(x = Day)) +
  geom_line(aes(y = ForecastedVar_NoIndicators, color = "No Indicators"), size = 1.2) +
  geom_line(aes(y = ForecastedVar_Indicators, color = "With Indicators"), size = 1.2) +
  geom_line(aes(y = RealizedVar, color = "Realized Volatility"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("No Indicators" = "blue",
                                "With Indicators" = "darkgreen",
                                "Realized Volatility" = "red")) +
  labs(
    title = "Rolling 1-Day GARCH Forecasts: With and Without Event Indicators",
    y = "Variance", x = "Forecast Day",
    color = "Legend"
  ) +
  theme_minimal()
```







## Key Event Modelling using GARCH
Estimate a model with 2025 as the test set and everything before as the training set.
Additionally, modelling some key events (COVID, first TRUMP term) using inidcators as key events.
```{r}
# Using data DGS10
train_start <- "2010-01-01"
train_end <- "2025-04-30"
train_data <- DGS10_diff[paste0(train_start, "/", train_end)]
colnames(train_data) <- "Vola"
test_data <- na.omit(DGS10_diff["2025-05-01/"])
spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1, 1)),
  mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
  distribution.model = "std"
)
fit_no_indicators <- ugarchfit(spec, train_data)
# Extract residuals and squared residuals
residuals_train <- residuals(fit_no_indicators, standardize = TRUE)
squared_residuals_train <- residuals(fit_no_indicators, standardize = FALSE)^2

```

```{r}
# Create indicators for key events
key_events <- data.frame(
  COVID = ifelse(index(train_data) >= as.Date("2020-03-01"), 1, 0),
  Trump = ifelse(index(train_data) >= as.Date("2016-11-08") & index(train_data) <= as.Date("2020-01-20"), 1, 0),
  Energy = ifelse(index(train_data) >= as.Date("2022-01-01"), 1, 0)
)
# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2
```



```{r}
# Forecasting with the fitted model
forecast_indicators <- ugarchforecast(fit_with_indicators, n.ahead = nrow(test_data))
forecast_no_indicators <- ugarchforecast(fit_no_indicators, n.ahead = nrow(test_data))
# Extract forecasted volatility
forecasted_volatility_indicators <- as.numeric(forecast_indicators@forecast$sigmaFor)
forecasted_volatility_no_indicators <- as.numeric(forecast_no_indicators@forecast$sigmaFor)
# Prepare data for plotting
forecasted_data <- data.frame(
  Date = index(test_data),
  ForecastedVolatility = forecasted_volatility_no_indicators,
  ForecastedVolatilityWithIndicators = forecasted_volatility_indicators
)
```

```{r}
coef(fit_no_indicators)
infocriteria(fit_no_indicators)

coef(fit_with_indicators)
infocriteria(fit_with_indicators)
```




```{r}
# Plotting the forecasted volatility along with the actual volatility for both the original model and the key event model
ggplot() +
  geom_line(data = forecasted_data, aes(x = Date, y = ForecastedVolatility), color = "blue", size = 1.2) +
  geom_line(data = forecasted_data, aes(x = Date, y = ForecastedVolatilityWithIndicators), color = "green", size = 1.2) +
  geom_line(data = data.frame(Date = index(test_data), ActualVolatility = abs(test_data)), aes(x = Date, y = ActualVolatility), color = "red", linetype = "solid", size = 1) +
  labs(title = "Forecasted Volatility with Key Events vs Actual Volatility",
       y = "Volatility (abs return or sigma)", x = "Date") +
  theme_minimal()

```

### Only the spikes of the events

```{r}
key_events <- data.frame(
  COVID = ifelse(index(train_data) >= as.Date("2020-01-01") & index(train_data) <= as.Date("2020-05-01"), 1, 0),
  Trump = ifelse(index(train_data) >= as.Date("2022-03-01") & index(train_data) <= as.Date("2022-10-01"), 1, 0),
  Energy = ifelse(index(train_data) >= as.Date("2023-12-01") & index(train_data) <= as.Date("2024-06-15"),1, 0)
)


# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2
```

```{r}
fit_with_indicators
```


## Break of structure - Bai-Perron test

```{r}
# Construct proper data frame with squared returns
squared_df <- data.frame(squared = DGS10_diff^2)

# Apply Bai–Perron test with intercept-only model
break_model <- breakpoints(squared ~ 1, data = squared_df, h = 0.15)

# Summarize breakpoints
summary(break_model)
```


```{r}
# Construct proper data frame with squared returns
squared_df <- data.frame(squared = DGS10_diff^2)

# Convert to zoo object
squared_zoo <- zoo(squared_df$squared, order.by = as.Date(rownames(squared_df)))

# Fit Bai–Perron model
break_model <- breakpoints(squared ~ 1, data = squared_df)

# Get breakpoint indices and convert to dates
bp_indices <- break_model$breakpoints
bp_dates <- as.Date(rownames(squared_df)[bp_indices])
```

```{r}
# Plot with actual dates
plot(squared_zoo, type = "l", 
     main = "Squared Returns with Bai–Perron Breakpoints",
     xlab = "Date", ylab = "Squared Returns")

# Add vertical lines at breakpoints
abline(v = bp_dates, col = "red", lty = 2)

print(bp_dates)
```

Two breakpoints: 2012-04-24 and 2022-02-03

### New events based on the breakpoints
```{r}
# Setup
N <- length(DGS10_diff)
forecast_horizon <- 63
full_index <- index(DGS10_diff)
realized_returns <- tail(DGS10_diff, forecast_horizon)
realized_var <- realized_returns^2

# Define rolling predictions container
forecasted_var_with_indicators <- numeric(forecast_horizon)

# Create full event matrix matching the entire series
event_indicators_full <- data.frame(
  Regime1 = ifelse(full_index >= as.Date("2012-04-24") & full_index <= as.Date("2012-10-01"), 1, 0),
  Regime2 = ifelse(full_index >= as.Date("2022-02-03") & full_index <= as.Date("2022-12-01"), 1, 0)
)

# Rolling forecast with external regressors
for (i in 1:forecast_horizon) {
  
  # Subset data up to time t + i - 1
  train_slice <- DGS10_diff[1:(N - forecast_horizon + i - 1)]
  reg_slice <- event_indicators_full[1:(N - forecast_horizon + i - 1), ]

  # Take the matching regressors for variance equation
  spec_step <- ugarchspec(
    variance.model = list(model = "sGARCH", garchOrder = c(1, 1),
                          external.regressors = as.matrix(reg_slice)),
    mean.model = list(armaOrder = c(0, 0), include.mean = TRUE),
    distribution.model = "std"
  )

  # Fit and forecast
  fit_step <- ugarchfit(spec_step, train_slice, solver = "hybrid", solver.control = list(trace = 0))
  fc <- ugarchforecast(fit_step, n.ahead = 1)
  forecasted_var_with_indicators[i] <- sigma(fc)^2
}
```


```{r}
# Comparison dataframe
comparison_df <- data.frame(
  Day = 1:forecast_horizon,
  RealizedVar = realized_var,
  ForecastedVar_NoIndicators = forecasted_var,  # from ugarchroll
  ForecastedVar_Indicators = forecasted_var_with_indicators
)

ggplot(comparison_df, aes(x = Day)) +
  geom_line(aes(y = ForecastedVar_NoIndicators, color = "No Indicators"), size = 1.2) +
  geom_line(aes(y = ForecastedVar_Indicators, color = "With Indicators"), size = 1.2) +
  geom_line(aes(y = RealizedVar, color = "Realized Volatility"), linetype = "dashed", size = 1.2) +
  scale_color_manual(values = c("No Indicators" = "blue",
                                "With Indicators" = "darkgreen",
                                "Realized Volatility" = "red")) +
  labs(
    title = "Rolling 1-Day GARCH Forecasts: With and Without Event Indicators",
    y = "Variance", x = "Forecast Day",
    color = "Legend"
  ) +
  theme_minimal()
```

Now, there is a difference between the blue and the green curves. The blue (no indicators) is usually above the green (with indicators).

```{r}
key_events <- data.frame(
  Regime1 = ifelse(full_index >= as.Date("2012-04-24") & full_index <= as.Date("2012-10-01"), 1, 0),
  Regime2 = ifelse(full_index >= as.Date("2022-02-03") & full_index <= as.Date("2022-12-01"), 1, 0)
)


# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2

fit_with_indicators@fit$matcoef
```

The volatility regime following February 3, 2022, which includes the early phases of the Russia–Ukraine war and subsequent economic instability, showed a statistically significant effect in the GARCH model's variance equation - it has a p-value of only 0.03613.

The earlier 2012 breakpoint likely reflects a shift to a lower-volatility environment, but without a statistically distinct impact relative to the base regime once GARCH dynamics are accounted for.


```{r}
plot(squared_zoo, type = "l", 
     main = "Squared Returns with Custom Dates",
     xlab = "Date", ylab = "Squared Returns")

# Add vertical lines at custom dates
custom_dates <- as.Date(c("2012-04-24", "2012-10-01", "2022-02-03", "2022-12-01"))
abline(v = custom_dates, col = "red", lty = 2)

```

### Trying to get a significant variable for COVID

```{r}
plot(squared_zoo, type = "l", 
     main = "Squared Returns with Custom Dates",
     xlab = "Date", ylab = "Squared Returns")

# Add vertical lines at custom dates
custom_dates <- as.Date(c("2020-02-01", "2020-04-01", "2022-02-03", "2023-01-01"))
abline(v = custom_dates, col = "red", lty = 2)
```

#### Both COVID and Energy Crisis

```{r}
key_events <- data.frame(
  COVID = ifelse(full_index >= as.Date("2020-01-20") & full_index <= as.Date("2020-03-20"), 1, 0),
  Energy = ifelse(full_index >= as.Date("2022-02-03") & full_index <= as.Date("2022-12-01"), 1, 0)
)


# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2

fit_with_indicators@fit$matcoef
```
Best indicator for COVID: 2020-01-20 - 2020-03-20. This is almost significant (on the 5% significance level) if we use it together with the other indicator variable.
Best indicator for energy crisis: 2022-02-03 - 2022-12-01

#### Only COVID
```{r}
key_events <- data.frame(
  COVID = ifelse(full_index >= as.Date("2020-01-20") & full_index <= as.Date("2020-03-20"), 1, 0)
)


# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2

fit_with_indicators@fit$matcoef
```

#### Only the Energy Crisis
```{r}
key_events <- data.frame(
  Energy = ifelse(full_index >= as.Date("2022-02-03") & full_index <= as.Date("2022-12-01"), 1, 0)
)


# Fit GARCH model with indicators
spec_with_indicators <- ugarchspec(
  variance.model = list(
    model = "sGARCH",
    garchOrder = c(1, 1),
    external.regressors = as.matrix(key_events)),
  mean.model = list(
    armaOrder = c(0, 0), 
    include.mean = TRUE,
    external.regressors = as.matrix(key_events)
  ),
  distribution.model = "std"
)
fit_with_indicators <- ugarchfit(spec_with_indicators, train_data$Vola)
# Extract residuals and squared residuals with indicators
residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = TRUE)
squared_residuals_train_with_indicators <- residuals(fit_with_indicators, standardize = FALSE)^2

fit_with_indicators@fit$matcoef
```

Interestingly, if we include both the COVID and the Energy Crisis indicators, then we get a better fit (smaller p-values for both variables) than if we use them alone.





